### 2025-04-28 Paper/beta-reducible-langs.tex
Added about beta reducability

\documentclass{llncs}

%\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, mathtools}
\usepackage{enumitem}
\usepackage{float}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{graphicx}
%\usepackage{mdframed}
%\usepackage{cleveref}

\def\labelitemi{$\top$}
\def\labelenumi{(\textbf{\arabic{enumi}})}

\usepackage{tikz}
\usetikzlibrary{trees}

\newcounter{algo}

\floatstyle{ruled}
\newfloat{algo}{h}{aux}
\floatname{algo}{Algorithm}

\input preamble

%\newcommand{\yoni}[1]{{\begin{mdframed}[linecolor=red]{\color{red}#1}\end{mdframed}}}
%\newcommand{\ari}[1]{{\begin{mdframed}[linecolor=blue]{\color{blue}#1}\end{mdframed}}}
\def\yoni#1{\par\noindent\fbox{\color{red}#1}\par}
\def\ari#1{\par\noindent\fbox{\color{blue}#1}\par}
\begin{document}

\title{Beta-Reduction}
\author{Ari Feiglin\inst{1}\and Yoni Zohar\inst{1}}
\institute{Bar-Ilan University, Ramat Gan, Israel}

\maketitle

\begin{abstract}

    In this paper we discuss an alternative approach to parsing context-free languages.
    \keywords{Context-Free Languages\and Parsing}

\end{abstract}

\section{$\beta$-Reducability}

In this section we investigate the formal language-theoretical properties of $\beta$-reduction.
In particular we discuss what sorts of languages can be $\beta$-reduced.
The definition of $\beta$-reduction in this section differs slightly from our previous definition, but the core principal
is preserved.
In particular, we will forgo priorities, which as we will see are beneficial implementation-wise, but unnecessary theoretically.

\subsection{The $\beta$-Reduction Hierarchy}

We begin by defining multiple classes of languages.
These are all special cases of the class we will begin to define now.

\begin{definition}

    Let $n>0$ be an integer, then an \emph{$n$-nondeterministic $\beta$-reduction schema} (shortened to $n-\NBR$ schema) is
    a tuple $B=(\Sigma,\Gamma,\square,\top,\hb)$ where
    \begin{enumerate}
        \item $\Sigma$ is a finite alphabet of terminals;
        \item $\Gamma$ is a finite alphabet of variables (we will assume that $\Sigma\subseteq\Gamma$, or that there exists a
            clear mapping of $\Sigma$ into $\Gamma$);
        \item $\square$ is a symbol in $\Gamma$ (the string beginning-marker);
        \item $\top$ is a symbol (potentially in $\Gamma$), or the empty word;
        \item $\hb$ is an \emph{initial $\beta$-reducer}, which is a function
        $$ \hb\colon\Gamma\times(\Gamma\cup\set\epsilon)\to\P((\Gamma\cup\set{\top})^{\leq n}\cup\set\perp) $$
            Where $\perp$ is a symbol distinct from any other previously discussed symbol.
    \end{enumerate}

\end{definition}

Let us focus on the initial $\beta$-reducer.
It takes as input two symbols and outputs a set of strings in $\Gamma\cup\set\top$ of length $\leq n$, and the set may also
contain the symbol $\perp$.
$\top$ is the target string, if we can reduce the string to $\top$ we consider the string part of the language.
And $\perp$ says that you can non-deterministically skip the reduction.

So for example if
$$ \hb(a,b) = \set{c,d,\perp},\quad \hb(b,a) = \set{c} $$
the string $aba$ can be reduced (in a single step) to one of the following:
$$ ca\quad da\quad ac $$
The first two result from reducing the first two characters ($a,b$) to one of their potential products, the third results from
skipping the first reduction (since $\perp\in\hb(a,b)$) and reducing the second two characters ($b,a$).

Let's formally define this process of reduction.
We assume that $\hb$ never returns an empty set (if we want to say that we can't reduce two symbols, we can return $\set\perp$).
Furthermore, we define concatenation of two sets of strings component-wise:
$A\circ B=\set{\xi_1\circ\xi_2}[\xi_1\in A,\xi_2\in B]$.
Where $\circ$ is the concatenation operator, which we will from now on leave out writing explicitly.
Furthermore, concatenation between a string and a set is done in the obvious way: $\xi A=\set\xi A$ and $A\xi=A\set\xi$.
Note that $\varnothing A=A\varnothing=\varnothing$, while $\epsilon A=A\epsilon=A$.

\begin{definition}

    Let $B$ be an $n$-\NBR, then we define its $\beta$-reduction function as follows.
    The function is $\beta\colon(\Gamma\cup\set\top)^*\to\P_{<\omega}((\Gamma\cup\set\top)^*)$ and defined as follows:
    \begin{enumerate}
        \item $\beta(\top)=\set\top$;
        \item $\beta(\sigma)=\hb(\sigma,\epsilon)-\set\perp$ if non-empty, and otherwise $\set\sigma$;
        \item For a string $\sigma_1\sigma_2\xi$, we first add
        $\bigl((\hb(\sigma_1,\sigma_2)-\set\perp)\xi\bigr)\cup\bigl((\hb(\sigma_1,\epsilon)-\set\perp)\sigma_2\xi\bigr)$ to
        the result.
        Then if $\perp\in\hb(\sigma_1,\sigma_2)\cup\hb(\sigma_1,\epsilon)$ we also add $\sigma_1\beta(\sigma_2\xi)$.
    \end{enumerate}
    This extends to a function $\beta\colon\P_{<\omega}((\Gamma\cup\set\top)^*)\to\P_{<\omega}((\Gamma\cup\set\top)^*)$
    in the natural way.
    We then define the relation $\xi\to\xi'$ to mean that $\xi'\in\beta(\xi)$; and we define $\sto$ to be the reflexive and
    transitive closure of $\to$.

    We then define the language of $B$ (or the language \emph{recognized} by $B$) to be
    $$ \L(B) = \set{w\in\Sigma^*}[\square w\sto\top] $$
    We define the class $\nNBR$ to be the class of all languages recognized by $\nNBR$ schemas.
    And we define $\sNBR=\bigcup_n\nNBR$.

\end{definition}

Note that an $\sNBR$-schema may be defined similar to an $\nNBR$-schema, but where the initial $\beta$-reducer returns finite
subsets of $(\Gamma\cup\set\top)^*$ (denoted $\P_{<\omega}((\Gamma\cup\set\top)^*)$).
This is since $\hb$ will return a longest string of some length $n$, and then the schema is an $\nNBR$ schema.

\subsection{Context-Free Languages}

Recall the definition of context-free languages:

\begin{definition}

    A \emph{context-free grammar} is a tuple $G=(\Sigma,V,S,P)$ where
    \begin{enumerate}
        \item $\Sigma$ is a finite alphabet of terminals;
        \item $V$ is a finite alphabet of variables;
        \item $S\in V$ is the initial variable;
        \item $P\subseteq V\times(\Sigma\cup V)^*$ is a set of production rules, where we write $A\vdash_G\alpha$ in place of
        $(A,\alpha)\in P$ (and we may omit the subscript from $\vdash$ when $G$ is understood).
    \end{enumerate}
    We extend the definition of $\vdash$ to a relation $(\Sigma\cup V)^*\times(\Sigma\cup V)^*$, where if
    $\alpha A\beta$ is a string of terminals and variables, $A\in V$ such that $A\vdash\gamma$, then
    $\alpha A\beta\vdash\alpha\gamma\beta$.
    $\vdash^*$ is the reflexive and transitive closure of $\vdash$.
    The language of $G$ is defined to be
    $$ \L(G) = \set{w\in\Sigma^*}[S\vdash^*w] $$
    A language is \emph{context-free} if it is the language of some context-free grammar.
    The class of context-free languages is denoted $\CFL$.

\end{definition}

Our goal of this section is to investigate the relationship between the classes $\CFL$ and $\nNBR,\sNBR$.

\begin{lemma}

    $\CFL\subseteq\sNBR$.

\end{lemma}

\begin{proof}

    Let $G=(\Sigma,V,S,P)$ be a context-free grammar.
    We construct an \sNBR-schema $B=(\Sigma,\Gamma,S,\epsilon,\hb)$, where $\Gamma=V\cup\Sigma\cup\overline\Sigma$ where
    $\overline\Sigma$ is a disjoint copy of $\Sigma$.
    For each production rule $A\vdash\alpha\in P$ we place $\perp,\overline\alpha^r\in\hb(A,\epsilon)$ (where $\overline\alpha$
    is $\alpha$ where $\sigma\in\Sigma$ is replaced with $\overline\sigma\in\overline\Sigma$).
    And for $\sigma\in\Sigma$, $\hb(\overline\sigma,\sigma)=\set{\epsilon}$.

    Note that $A\to^*\overline\alpha\in(V\cup\overline\Sigma)^*$, and so when reducing $A$, we can choose any variable to
    reduce.
    That is, if $A_1\to\alpha_1$ and $A_2\to\alpha_2$, and $A\to^*\gamma_1A_1\gamma_2A_2\gamma_3$, then we have that in a single
    reduction we can either get $\gamma_1\alpha_1\gamma_2A_2\gamma_3$ or we get  $\gamma_1A_1\gamma_2\alpha_2\gamma_3$.
    This is because elements of $\overline\Sigma$ won't be reduced, and all other reductions (of variables) may be skipped, as
    they contain $\perp$.

    Now we claim that for any $n\geq0$ and $A\in V$, if $A\vdash^n\alpha$ iff $A\to^n\overline\alpha^r$.
    This is clear by induction: for $n=0$ this is because $\alpha=A$.
    For $n=1$ this is by definition.
    Now suppose $A\vdash^n\alpha_1B\alpha_3$ and $B\vdash\alpha_2$, so that $A\vdash^{n+1}\alpha_1\alpha_2\alpha_3$.
    Then by induction $A\to^n\alpha_3^rB\alpha_1^r$ and by definition $B\to\alpha_2^r$, so that we have
    $A\to^{n+1}\alpha_3^r\alpha_2^r\alpha_1^r=(\alpha_1\alpha_2\alpha_3)^r$ as required.
    The other direction is proven in a similar way.

    If $w\in\L(G)$, we have that $S\vdash^*w$, and so $Sw\sto\overline w^rw$, which clearly then $\sto\epsilon$.
    So $w\in\L(B)$.

    Now if $Aw\sto\overline\alpha w_2$, then we claim there exists a $w_1$ such that $w=w_1w_2$ and $A\vdash^*w_1\alpha^r$.
    We do this by induction on the number of steps in the reduction, $n$.
    For $n=0,1$ this is clear ($w_1=\epsilon$) in both cases.
    Now suppose $Aw\to^n\overline\alpha w_2$, then $w=w_1w_2$ and $A\vdash^*w_1\alpha^r$.
    If the next reduction is in $\overline\alpha$, this must be a variable reduction, which means that $A\vdash^*w_1(\alpha')^r$
    and $Aw\to^{n+1}\overline\alpha' w_2$, so $w'_1=w_1$ still satisfies the requirement.
    Otherwise, the reduction is of the form
    $$ Aw\to^n\overline\alpha'\overline\sigma\sigma w'_2 \to\overline\alpha'w'_2 $$
    In which case $w=w_1w_2=w_1\sigma w'_2$, so that $w'_1$ must be $w_1\sigma$.
    Now, notice that $A\vdash^*w_1\alpha^r=w_1(\alpha'\sigma)^r=w_1\sigma(\alpha')^r=w'_1(\alpha')^r$ as required.

    So we have that $Sw\sto\epsilon$, so $w_2=\epsilon$ and thus $w_1=w$, meaning $S\vdash^*w$.
    Thus if $w\in\L(B)$, $w\in\L(G)$, as required.
    \qed

\end{proof}

Notice that the parameter of the $\sNBR$-schema is dependent only on the length of the production strings in $G$.
That is $B$ is an $\nNBR$-schema where $n=\max\set{\abs\alpha}[A\vdash_G\alpha]$.
Every context-free grammar has an equivalent in Chomsky normal form (CNF), whose $n$ is $2$.
Thus $\CFL\subseteq\pNBR2$.

In fact, $\sNBR=\pNBR2$; this is not difficult to prove.
Take an $\nNBR$-schema $B=(\Sigma,\Gamma,\square,\top,\hb)$ and define a $\pNBR2$-schema
$B_2=(\Sigma,\Gamma_2,\square,\top,\hb_2)$ as follows: $\Gamma_2$ will contain $\Gamma$ as well as every prefix of strings
returned by $\hb$ (each prefix is distinct, if two strings have the same prefix, or the same string is returned by two different
reductions, we consider them distinct).
Given such a prefix $\xi$, denote the symbol in $\Gamma_2$ by $\mbox\xi$.
Then if $\sigma_1\cdots\sigma_n\in\hb(\tau_1,\tau_2)$, we define
$$ \hb_2(\tau_1,\tau_2) = \mbox{\sigma_1\cdots\sigma_n},\quad
\hb_2(\mbox{\sigma_1\cdots\sigma_k},\epsilon)=\set{\mbox{\sigma_1\cdots\sigma_{k-1}}\sigma_k}\hbox{ for $k\leq n$} $$
If $\perp\hb(\tau_1,\tau_2)$, then $\perp\in\hb_2(\tau_1,\tau_2)$ as well.

\begin{lemma}

    $\sNBR=\pNBR2$

\end{lemma}

Though, the most important result is as follows:

\begin{lemma}

    $\CFL\subseteq\pNBR1$

\end{lemma}

\begin{proof}

    Let $G$ be a context-free grammar in Chomsky normal form.
    We define $B=(\Sigma,\Gamma,\square,\epsilon,\hb)$, where $\Gamma=\Sigma\cup V\cup\set\square$ and for every rule
    $A\vdash_GXY$, we add the reduction $X,Y\to_BA$.
    And for every rule $A\vdash_G\sigma$, we add the reduction $\sigma,\epsilon\to_BA$.
    Then for every two symbols $X,Y$ we add the reduction $X,Y\to_B\vdash$.
    If $S\vdash_G\epsilon$ is a rule, then we also add the reduction $\square,\epsilon\to_B\epsilon$.
    We also add the rule $\square,S\to_B\epsilon$.

    Now, we claim that for $\alpha,\gamma\in\Gamma^*$, $\alpha\vdash\gamma$ if and only if $\gamma\to\alpha$.
    Indeed, if $\alpha=\alpha_1A\alpha_2$ and we use the rule $A\to XY$, then $\gamma=\alpha_1XY\alpha_2$.
    And we can use the reduction $X,Y\to A$ to get $\gamma\to\alpha_1A\alpha_2=\alpha$ (since every reduction contains $\perp$,
    we can arbitrarily skip strings).
    And if $\alpha=\alpha_1A\alpha_2$ and we use the rule $A\to\sigma$, then $\gamma=\alpha_1\sigma\alpha_2$ and of course we
    have $\gamma\to\alpha$.

    Proving the other direction is similar.
    Thus we have $\alpha\vdash\gamma$ if and only if $\gamma\to\alpha$.
    Thus $\alpha\vdash^*\gamma$ if and only if $\gamma\to^*\alpha$.
    In particular $S\vdash^*\gamma$ (i.e. $\gamma\in\L(G)$) if and only if $\gamma\to^*S$.
    This means that $\square\gamma\to^*\square S\to\epsilon$, so $\gamma\in\L(B)$.
    If $\square$ is reduced before $\gamma$ gets to $S$, we will have the case $\square\gamma\to^*S$, which won't reduce.
    So in either case we have that $S\vdash^*\gamma\iff\square\gamma\to^*S$.
    This works for $\gamma=\epsilon$ as well.
    So we have $\L(G)=\L(B)$ as required.
    \qed

\end{proof}

\subsection{Deterministic $\beta$-Reduction}

\begin{definition}

    An $\nNBR$-schema $B$ is \emph{deterministic} if for every $\sigma\in\Gamma,\tau\in\Gamma\cup\set\epsilon$, we have that
    $\lvert\hb(\sigma,\tau)\rvert=1$.
    That is $\hb(\sigma,\tau)=\set\alpha$ or $\hb(\sigma,\tau)=\set\perp$.
    We call a deterministic $\nNBR$-schema an $\nBR$-schema.
    Equivalently an $\nNBR$-schema is one where $\hb$ is a partial function
    $$ \hb\colon\Gamma\times(\Gamma\cup\set\epsilon) \partial (\Gamma\cup\set\top)^{\leq n} $$
    (Where an image not being defined is equivalent to defining it to be $\set\perp$.)

    The class of languages accepted by $\nBR$-schemas is denoted $\nBR$.
    Let $\sBR=\bigcup_{n>0}\nBR$.

\end{definition}

Our goal in this section is to understand the relationship between $\nBR$ and the class of deterministic context-free languages.
We denote this class by $\DCFL$.
Recall that a deterministic context-free language is one that can be accepted by a deterministic pushdown automaton.
Our previous proof techniques do not work for the deterministic case, as they rely on the ability to non-deterministically
choose whether or not to reduce.

Though notice that our proof that $\sNBR=\pNBR2$ converts deterministic schemas to deterministic schemas.
So we immediately have that

\begin{lemma}

    $\sBR=\pBR2$

\end{lemma}

We will use an equivalent definition of deterministic context-free languages to show that $\DCFL\subseteq\pBR2$.

\begin{definition}

    An \emph{LR parser} is a tuple $P=(\Sigma,Q,V,\action,\goto,q_0,\square)$ where
    \begin{enumerate}
        \item $\Sigma$ is a finite alphabet of terminals;
        \item $Q$ is a finite set of states (disjoint from $\Sigma$);
        \item $V$ is a finite set of variables (disjoint from $\Sigma$ and $Q$);
        \item $\action$ is a partial function
        \begin{multline*}
            \action\colon Q\times\Sigma\cup\set\square\partial\\
            \set{\shift(q)}[q\in Q]\cup\set{\reduce(A,n)}[A\in V,n\in{\bb Z}_{\geq0}]\cup\set\accept
        \end{multline*}
        ($\shift,\reduce,\accept$ are simply distinct symbols.)
        \item $\goto$ is a partial function
        $$ \goto\colon Q\times V \partial Q $$
        \item $q_0\in Q$ is the initial state;
        \item $\square$ is a symbol not in $\Sigma\cup Q\cup V$, the string endmarker.
    \end{enumerate}

\end{definition}

\begin{definition}

    A \emph{configuration} of an LR parser $P$ is a pair $c\in Q^*\times(\Sigma\cup\set\square)^*$, i.e. a pair
    of the form $(p_0\cdots p_n\;;\;\sigma_1\cdots\sigma_m)$.
    We define $P(c)$ as follows: suppose $c=(q_0\cdots q_m\;;\;\sigma_1\cdots\sigma_n\square)$, then
    \begin{enumerate}
        \item if $\action(q_m,\sigma_1)=\shift(p)$ then
        $$ P(c) = (q_0\cdots q_mp\;;\;\sigma_2\cdots\sigma_n\square) $$
        \item if $\action(q_m,\sigma_1)=\reduce(A,k)$ then let $p=\goto(q_{m-k},A)$ and
        $$ P(c) = (q_0\cdots q_{m-k}p\;;\;\sigma_1\cdots\sigma_n\square) $$
        if $\goto(p_{m-k},A)$ is undefined, $P(c)=\bot$ (a special configuration invariant under $P$);
        \item if $\action(q_m,\sigma_1)=\accept$ then $P(c)=\top$ (also invariant under $P$);
        \item if $\action(q_m,\sigma_1)$ is undefined then $P(c)=\bot$.
    \end{enumerate}
    The language of $P$ is
    $$ \L(P) = \set{w\in\Sigma^*}[\hbox{there exists an $n>0$ such that $P^n(q_0\;;\;w\square)=\top$}] $$

\end{definition}

A well-known theorem is the following:

\begin{theorem}

    A language $\L$ is deterministic context-free if and only if there exists an LR parser such that $\L(P)=\L$.

\end{theorem}

\begin{lemma}

    $\DCFL\subseteq\sBR$

\end{lemma}

\begin{proof}

    Let $P$ be an LR parser.
    We define a $\sBR$-schema $(\Sigma,\Gamma,q_0,\top,\hb)$ where $\Gamma$ consists of $\Sigma\cup Q$ and for every
    $\reduce(A,n)$ in the image of $\action$, $M_{k,A}$ for $k\leq n$ and $A\in V$, and we add $\perp,\top$ as well.

    For every $q\in Q$ and $\sigma\in\Sigma$,
    \begin{enumerate}
        \item if $\action(q,\sigma)=\shift(p)$ then $\hb(q,\sigma)=qp$;
        \item if $\action(q,\sigma)=\reduce(A,n)$, then $\hb(q,\sigma)=qM_{n,A}\sigma$;
        \item if $\action(q,\sigma)=\accept$, then $\hb(q,\sigma)=\top$;
        \item if $\action(q,\sigma)$ is undefined, then $\hb(q,\sigma)=\bot$;
        \item $\hb(q,M_{n,A})=M_{n-1,A}$ if $n>0$;
        \item $\hb(q,M_{0,A}=q\goto(q,A)$ if defined, otherwise $\perp$;
        \item $\hb(\sigma,\top)=\hb(\top,\sigma)=\top$.
    \end{enumerate}
    We also add $\hb(q,\epsilon)$ according to $\action(q,\square)$.
    We can assume that this is not a $\shift$ operation.

    It's not hard to see that for a given configuration $c=(v;w)$, we have that $P(c)=(v',w')$ then $vw\to^* v'w'$.
    And if $vw\to^*v'w'$ then $P^k(c)=(v',w')$ for some $k\geq0$.
    Thus we get $\L(P)=\L(B)$ as required.
    \qed

\end{proof}

Since $\sBR=\pBR2$, we get

\begin{lemma}

    $\DCFL\subseteq\pBR2$

\end{lemma}

\end{document}